\textcolor{Navy}{\textsc{Convexity adjustments with a bit of Malliavin}}

### 2022-10-25 Latex/paper.tex
# cms covexity adjustment (not end)
# fix several bugs in cms example

\subsection{CMSs}
We will define the swap rate from $T_a$ to $T_b$ at time $t$
\begin{equation*}
S_{a,b} = \frac{\sum_{i=1}^{n_E}\delta_{t^{E}_{i-1}, t^{E}_i} L^{E}(t,t^{E}_{i-1}, t^{E}_{i}) P_{ois}(t,t^{E}_{i})}{01(t,T_a,T_b)}
\end{equation*}
where
\begin{align*}
01(t,T_a,T_b) = \sum_{j=1}^{n_f} \delta_{t^{f}_{i-1}, t^{f}_i} P_{ois}(t,t^{f}_{j}) \\
T_a=T^{E}_0 < T^{E}_i< \cdots < T^{E}_{n_E}=T_b \quad i=0,\cdots,n_E&  \\
T_a=T^{f}_0 < T^{f}_j< \cdots < T^{f}_{n_f}=T_b \quad j=0,\cdots,n_f&
\end{align*}
We suppose, that we have a flow in $T_a < T_p < T_b$ with value $S_{a,b}(T_a)$. Then, the present value of this flow
\begin{align}\label{cms_expedted_value}
P_{ois}(0,T_a)\mathbb{E}^{\mathbb{Q}^{T_{a}}}\left(S_{a,b}(T_a) P_{ois}(T_a,T_p) \right) &= 01(0,T_a,T_b) \mathbb{E}^{\mathbb{Q}^{01(a,b)}}\left(S_{a,b}(T_a) \frac{P_{ois}(T_a,T_p)}{01(T_a,T_a,T_b) }\right) \nonumber \\
&= 01(0,T_a,T_b) \mathbb{E}^{\mathbb{Q}^{01(a,b)}}\left(S_{a,b}(T_a) M(T_a,T_p)\right)
\end{align}
where $M(t, T_p)= \frac{P_{ois}(t,T_p) }{01(t,T_a,T_b)} $. In order to reduce the complexity to explain the movement od the forward curve, it is a common pratice to asumme that
\begin{equation*}
M(t) = f(S_{a,b}(t))
\end{equation*}
where $f(\cdot)$ is known as mapping function. There is a extensive literature how to define this function (see Piterbarg Volume III), in later we will see some particular examples. Under the last assumption and applying Clark-Ocone to $f(S_{a,b}(t))$ we get
\begin{align*}
f(S_{a,b}(T_a)) &= \mathbb{E}^{\mathbb{Q}^{01(a,b)}}\left(f(S_{a,b}(T_a))\right) + \int_{0}^{T_a} \mathbb{E}_s^{01(a,b)}\left( D_s (f(S_{a,b}(T_a))) \right) dW^{01(a,b)}_{s} \\
&= \frac{P_{ois}(0,T_p) }{01(0,T_a,T_b)} + \int_{0}^{T_a} \mathbb{E}_s^{01(a,b)}\left(f^{\prime}(S_{a,b}(T_a)) \partial_{x}S_{a,b}(T_a) D_s x_{T_a} \right) dW^{01(a,b)}_{s}. 
\end{align*}
Therefore, we can define the convexity adjustment for a CMS the next way
\begin{equation*}
CA_{CMS}(T_a,T_b) = 01(0,T_a,T_b) \mathbb{E}^{\mathbb{Q}^{01(a,b)}}\left(S_{a,b}(T_a) \int_{0}^{T_a} \mathbb{E}_s^{01(a,b)}\left( f^{\prime}(S_{a,b}(T_a)) \partial_{x}S_{a,b}(T_a) D_s x_{T_a} \right) dW^{01(a,b)}_{s} \right).
\end{equation*}
Now if we apply the adjoint of the Malliavin derivative (see Elisa-David) we get

\begin{align}
CA_{CMS}(T_a,T_b) &= 01(0,T_a,T_b) \mathbb{E}^{\mathbb{Q}^{01(a,b)}}\left( \int_{0}^{T_a} \partial_{x}S_{a,b}(T_a) D_s x_{T_a} \mathbb{E}_s^{01(a,b)}\left(f^{\prime}(S_{a,b}(T_a)) \partial_{x}S_{a,b}(T_a) D_s x_{T_a} \right) ds \right) \nonumber \\
&\approx 01(0,T_a,T_b)  \mathbb{E}^{\mathbb{Q}^{01(a,b)}}\left( \int_{0}^{T_a} \left(\partial_{x}S_{a,b}(0)\right)^{2} f^{\prime}(S_{a,b}(0))  D_s x_{T_a} \mathbb{E}_s^{01(a,b)}\left(D_s x_{T_a} \right) ds \right) \nonumber \\
& \approx 01(0,T_a,T_b)  \mathbb{E}^{\mathbb{Q}^{01(a,b)}}\left( \int_{0}^{T_a} \left(\partial_{x}S_{a,b}(0)\right)^{2} f^{\prime}(S_{a,b}(0)) \beta(s,T_a,x_0,\hat{y}_{s})  D_s x_{T_a}  ds \right)
\end{align}

\begin{remark}\label{drift_01_spot_measure}
Before to continue, we must to calculate the Girsanov's measure change from $\mathbb{Q}$ to $\mathbb{Q}^{01(a,b)}$. In order to achieve this target, we will use a continuous version of $01(t,T_a,T_b)$ i.e
\begin{equation*}
01^{c}(t,T_a,T_b) = \int_{T_a}^{T_b} P_{ois}(t,u) du.
\end{equation*}
It is easy to see that if we define $z_t=\frac{01^{c}(t,T_a,T_b)}{\beta_t}$ and apply Itô's formula we get
\begin{equation*}
dW^{01(a,b)}_t = dW^{\mathbb{Q}}_t + \lambda_{t}dt 
\end{equation*}
where $\lambda_t = \frac{\int_{T_a}^{T_b} P_{ois}(t,u) \nu(t,u) du}{01^{c}(t,T_a,T_b)}$. Then we can do the next approximation for $\lambda_t$
\begin{equation*}
\hat{\lambda}_t \approx \frac{\int_{T_a}^{T_b} P_{ois}(t,u,x_0,\hat{y}_t) \nu(t,u) du}{01^{c}(t,T_a,T_b,x_0,\hat{y}_t)} 
\end{equation*}
\end{remark}
Now, if we apply (\ref{drift_01_spot_measure}) we get that 
\begin{equation*}
\mathbb{E}^{\mathbb{Q}^{01(a,b)}}\left( D_s x_{T_a} \right) \approx \beta(s,T_a,x_0,\hat{y}_{s}) \exp\left(-\int_{s}^{T_a} \frac{\int_{T_a}^{T_b} P_{ois}(u,u^{\prime},x_0,\hat{y}_u) \nu(t,u^{\prime}) du^{\prime}}{01^{c}(u,T_a,T_b,x_0,\hat{y}_u)} \partial_x \beta(u,t_0,x_0,\hat{y}_u) du \right)
\end{equation*}
Therefore we have the next for the convexity adjustment

@@

\begin{equation}\label{cms_adjustment}
CA_{CMS}(T_a,T_b) \approx 01(0,T_a,T_b) \int_{0}^{T_a} \left(\partial_{x}S_{a,b}(0)\right)^{2} f^{\prime}(S_{a,b}(0)) \beta^{2}(s,T_a,x_0,\hat{y}_{u}) \Omega(s, T_a)    ds
\end{equation}
with 
$$\Omega(s, T_a) = \exp\left(-\int_{s}^{T_a} \frac{\int_{T_a}^{T_b} P_{ois}(u,u^{\prime},x_0,\hat{y}_u) \nu(t,u^{\prime}) du^{\prime}}{01^{c}(u,T_a,T_b,x_0,\hat{y}_u)} \partial_x \beta(u,t_0,x_0,\hat{y}_u) du \right) 
$$
}}
\begin{example}
In ghe case of a Hull-White model, we showed in the example  (\ref{example_convexity_hw_ois}) that 
$$
\partial_x \beta(u,t_0,x_0,\hat{y}_u)=0
$$
\end{example}
Therefore (\ref{cms_adjustment}) is reduced to
$$
01(0,T_a,T_b) \int_{0}^{T_a} \left(\partial_{x}S_{a,b}(0)\right)^{2} f^{\prime}(S_{a,b}(0)) \beta^{2}(s,T_a,x_0,\hat{y}_{u}) ds
$$

>>>>

\subsection{CMSs}
We will define the swap rate from $T_a$ to $T_b$ at time $t$
\begin{equation*}
S_{a,b}(t) = \frac{\sum_{i=1}^{n_E}\delta_{t^{E}_{i-1}, t^{E}_i} L^{E}(t,t^{E}_{i-1}, t^{E}_{i}) P_{ois}(t,t^{E}_{i})}{01(t,T_a,T_b)}
\end{equation*}
where
\begin{align*}
01(t,T_a,T_b) = \sum_{j=1}^{n_f} \delta_{t^{f}_{i-1}, t^{f}_i} P_{ois}(t,t^{f}_{j}) \\
T_a=T^{E}_0 < T^{E}_i< \cdots < T^{E}_{n_E}=T_b \quad i=0,\cdots,n_E&  \\
T_a=T^{f}_0 < T^{f}_j< \cdots < T^{f}_{n_f}=T_b \quad j=0,\cdots,n_f&
\end{align*}
The same way, let us define
\begin{equation*}
S^{ois}_{a,b}(t) = \frac{P_{ois}(t,T^{E}_a) - P_{ois}(t,T^{E}_b)}{01(t,T_a,T_b)}. 
\end{equation*}

We suppose, that we have a flow in $T_a < T_p < T_b$ with value $S_{a,b}(T_a)$. Then, the present value of this flow
\begin{align}\label{cms_expedted_value}
P_{ois}(0,T_a)\mathbb{E}^{\mathbb{Q}^{T_{a}}}\left(S_{a,b}(T_a) P_{ois}(T_a,T_p) \right) &= 01(0,T_a,T_b) \mathbb{E}^{\mathbb{Q}^{01(a,b)}}\left(S_{a,b}(T_a) \frac{P_{ois}(T_a,T_p)}{01(T_a,T_a,T_b) }\right) \nonumber \\
&= 01(0,T_a,T_b) \mathbb{E}^{\mathbb{Q}^{01(a,b)}}\left(S_{a,b}(T_a) M(T_a,T_p)\right)
\end{align}
where $M(t, T_p)= \frac{P_{ois}(t,T_p) }{01(t,T_a,T_b)} $. In order to reduce the complexity to explain the movement od the forward curve, it is a common pratice to asumme that
\begin{equation*}
M(t) = f(S^{ois}_{a,b}(t))
\end{equation*}
where $f(\cdot)$ is known as mapping function. There is a extensive literature how to define this function (see Piterbarg Volume III), in later we will see some particular examples. Under the last assumption and applying Clark-Ocone to $f(S_{a,b}(t))$ we get
\begin{align*}
f(S^{ois}_{a,b}(T_a)) &= \mathbb{E}^{\mathbb{Q}^{01(a,b)}}\left(f(S^{ois}_{a,b}(T_a)\right) + \int_{0}^{T_a} \mathbb{E}_s^{01(a,b)}\left( D_s (f(S^{ois}_{a,b}(T_a)) \right) dW^{01(a,b)}_{s} \\
&= \frac{P_{ois}(0,T_p) }{01(0,T_a,T_b)} + \int_{0}^{T_a} \mathbb{E}_s^{01(a,b)}\left(f^{\prime}(S^{ois}_{a,b}(T_a) \partial_{x}S^{ois}_{a,b}(T_a) D_s x_{T_a} \right) dW^{01(a,b)}_{s}. 
\end{align*}
Therefore, we can define the convexity adjustment for a CMS the next way
\begin{align*}
CA_{CMS}(T_a,T_b) = \\
01(0,T_a,T_b) \mathbb{E}^{\mathbb{Q}^{01(a,b)}}\left(S_{a,b}(T_a) \int_{0}^{T_a} \mathbb{E}_s^{01(a,b)}\left( f^{\prime}(S^{ois}_{a,b}(T_a)) \partial_{x}S^{ois}_{a,b}(T_a) D_s x_{T_a} \right) dW^{01(a,b)}_{s} \right)
\end{align*}
Now if we apply the adjoint operator of the Malliavin derivative (see Elisa-David) we get

\begin{align*}
&CA_{CMS}(T_a,T_b)  \\
&=01(0,T_a,T_b) \mathbb{E}^{\mathbb{Q}^{01(a,b)}}\left( \int_{0}^{T_a} \partial_{x}S_{a,b}(T_a) D_s x_{T_a} \mathbb{E}_s^{01(a,b)}\left(f^{\prime}(S^{ois}_{a,b}(T_a)) \partial_{x}S^{ois}_{a,b}(T_a) D_s x_{T_a} \right) ds \right) \nonumber \\
&\approx 01(0,T_a,T_b)  \mathbb{E}^{\mathbb{Q}^{01(a,b)}}\left( \int_{0}^{T_a} \partial_{x}S_{a,b}(T_a, x_0, y_0) \partial_{x}S^{ois}_{a,b}(T_a, x_0, y_0) f^{\prime}(S^{ois}_{a,b}(T_a, x_0, y_0)  D_s x_{T_a} \mathbb{E}_s^{01(a,b)}\left(D_s x_{T_a} \right) ds \right) \nonumber \\
& \approx 01(0,T_a,T_b)  \mathbb{E}^{\mathbb{Q}^{01(a,b)}}\left( \int_{0}^{T_a} \partial_{x}S_{a,b}(T_a, x_0, y_0) \partial_{x}S^{ois}_{a,b}(T_a, x_0, y_0) f^{\prime}(S^{ois}_{a,b}(T_a, x_0, y_0) \beta(s,T_a,x_0,\hat{y}_{s})  D_s x_{T_a}  ds \right)
\end{align*}

\begin{remark}\label{drift_01_spot_measure}
Before to continue, we must to calculate the Girsanov's measure change from $\mathbb{Q}$ to $\mathbb{Q}^{01(a,b)}$. In order to achieve this target, we will use a continuous version of $01(t,T_a,T_b)$ i.e
\begin{equation*}
01^{c}(t,T_a,T_b) = \int_{T_a}^{T_b} P_{ois}(t,u) du.
\end{equation*}
It is easy to see that if we define $z_t=\frac{01^{c}(t,T_a,T_b)}{\beta_t}$ and apply Itô's formula we get
\begin{equation*}
dW^{01(a,b)}_t = dW^{\mathbb{Q}}_t + \lambda_{t}dt 
\end{equation*}
where $\lambda_t = \frac{\int_{T_a}^{T_b} P_{ois}(t,u) \nu(t,u) du}{01^{c}(t,T_a,T_b)}$. Then we can do the next approximation for $\lambda_t$
\begin{equation*}
\hat{\lambda}_t \approx \frac{\int_{T_a}^{T_b} P_{ois}(t,u,x_0,\hat{y}_t) \nu(t,u) du}{01^{c}(t,T_a,T_b,x_0,\hat{y}_t)} 
\end{equation*}
\end{remark}
Now, if we apply (\ref{drift_01_spot_measure}) we get that 
\begin{equation*}
\mathbb{E}_s^{\mathbb{Q}^{01(a,b)}}\left( D_s x_{T_a} \right) \approx \beta(s,T_a,x_0,\hat{y}_{s}) \exp\left(-\int_{s}^{T_a} \frac{\int_{T_a}^{T_b} P_{ois}(u,u^{\prime},x_0,\hat{y}_u) \nu(t,u^{\prime}) du^{\prime}}{01^{c}(u,T_a,T_b,x_0,\hat{y}_u)} \partial_x \beta(u,t_0,x_0,\hat{y}_u) du \right)
\end{equation*}
Therefore we have the next for the convexity adjustment

@@

\begin{align}\label{cms_adjustment}
&CA_{CMS}(T_a,T_b) \nonumber \\
&\approx 01(0,T_a,T_b) \int_{0}^{T_a} \partial_{x}S_{a,b}(T_a, x_0, y_0) \partial_{x}S^{ois}_{a,b}(T_a, x_0, y_0)  f^{\prime}(S^{ois}_{a,b}(T_a, x_0, y_0)) \beta^{2}(s,T_a,x_0,\hat{y}_{s}) \Omega(s, T_a)    ds
\end{align}
with 
$$\Omega(s, T_a) = \exp\left(-2\int_{s}^{T_a} \frac{\int_{T_a}^{T_b} P_{ois}(u,u^{\prime},x_0,\hat{y}_u) \nu(t,u^{\prime}) du^{\prime}}{01^{c}(u,T_a,T_b,x_0,\hat{y}_u)} \partial_x \beta(u,t_0,x_0,\hat{y}_u) du \right) 
$$
}}
\begin{example}
In ghe case of a Hull-White model, we showed in the example  (\ref{example_convexity_hw_ois}) that 
$$
\partial_x \beta(u,t_0,x_0,\hat{y}_u)=0
$$
\end{example}
Therefore (\ref{cms_adjustment}) is reduced to
\begin{equation}\label{cms_general_convexity}
01(0,T_a,T_b) \int_{0}^{T_a} \partial_{x}S_{a,b}(T_a,x_0,y_0) \partial_{x}S^{ois}_{a,b}(T_a,x_0,y_0) f^{\prime}(S^{ois}_{a,b}(T_a,x_0,y_0)) \beta^{2}(s,T_a,x_0,\hat{y}_{s}) ds
\end{equation}
A common practice in market, it is suppose that $f(x)=\alpha x + \beta$ (TSR model). We must to impose two equations in order to get $\alpha$ and $\beta$. The first equation follows straightforward from the definition of $M(t,T_p)$. Since, $M(t,T_p)$ is a $01(a,b)-$martingale we must have that
\begin{equation*}
\mathbb{E}^{01(a,b)}\left(\alpha S^{ois}_{a,b}(T_a) + \beta \right) = \frac{P_{ois}(0,T_p)}{01(0,T_a,T_b)}.
\end{equation*}
Therefore
\begin{equation*}
\beta = \frac{P_ois(0,T_p)}{01(0,T_a,T_b,x_0,y_0)} - \alpha S^{ois}_{a,b}(0,x_0,y_0).
\end{equation*}
Under a separable HJM dynamic we must have that
\begin{equation*}
\alpha \partial_{x} S^{ois}_{a,b}(T_a,x_0,y_0) = - M(0,T_p)\left(G(T_a,T_p)+\frac{\partial_x 01(T_a,T_a,T_b,x_0,y_0)}{01(T_a,T_a,T_b,x_0,y_0)}\right).
\end{equation*}
Then
\begin{equation*}
\alpha = -\frac{M(T_a,T_p)\left(G(T_a,T_p)+\frac{\partial_x 01(T_a,T_a,T_b,x_0,y_0)}{01(T_a,T_a,T_b,x_0,y_0)}\right)}{\partial_{x} S^{ois}_{a,b}(T_a,x_0,y_0)}.
\end{equation*}
Then from the last equalities, we can approximate the convexity adjustment of a CMS rate as
\begin{align*}
CA_{CMS}(T_a,T_b)\\
\approx - P_{ois}(0,T_p,x_0,y_0) \left(G(T_a,T_p) +\frac{\partial_x 01(T_a,T_a,T_b,x_0,y_0)}{01(T_a,T_a,T_b,x_0,y_0)}\right) \partial_{x}S_{a,b}(0,x_0,y_0) \int_{0}^{T_a}  \beta^{2}(s,T_a,x_0,\hat{y}_{u}) ds
\end{align*}
In the next figure, we can see the convexity adjustment for a CMS where the tenor for the underlying swap is 5Y using the last approximation and Monte Carlo method
